# Test data {#test_data}

```{r test_data_eval, echo = FALSE, ref.label = "test_data_show", message = FALSE, results = "hide"}
```

We will benchmark the 
`r BiocStyle::Githubpkg("PeteHaitch/DelayedMatrixStats")` package using 
different types of matrix-like data:

- `dense_matrix`: A dense *matrix* with `r ncol(dense_matrix)` columns and `r nrow(dense_matrix)` rows (`r format(object.size(dense_matrix), units = "MB")`)
- `sparse_matrix`: A sparse *dgcMatrix* with `r ncol(sparse_matrix)` columns and `r nrow(sparse_matrix)` rows where `r round(100 * sum(sparse_matrix == 0) / length(sparse_matrix), 2)`% of entries are zero (`r format(object.size(sparse_matrix), units = "MB")`)
- `rle_matrix`: An run-length encoded column *RleMatrix* with `r ncol(rle_matrix)` columns and `r nrow(rle_matrix)` rows  (`r format(object.size(rle_matrix), units = "MB")`)

```{r test_data_show, echo = TRUE, message = FALSE}
library(DelayedMatrixStats)
library(Matrix)
library(HDF5Array)

# Dense matrix
dense_matrix <- DelayedArray(matrix(runif(20000 * 600), nrow = 20000,
                                  ncol = 600))
dense_matrix

# 60% zero elements
sparse_matrix <- seed(dense_matrix)
zero_idx <- sample(length(sparse_matrix), 0.6 * length(sparse_matrix))
sparse_matrix[zero_idx] <- 0
sparse_matrix <- DelayedArray(Matrix::Matrix(sparse_matrix, sparse = TRUE))
sparse_matrix

# HDF5-backed dense matrix
# hdf5_matrix <- as(dense_matrix, "HDF5Array")

# Run-length encoded column matrix
rle_matrix <- RleArray(Rle(sample(2L, 200000 * 6 / 10, replace = TRUE), 100),
                       dim = c(2000000, 6))
rle_matrix
```

Obviously, this is not a comprehensive set of inputs. Rather, it chosen to be 
somewhat representative of some typical genomics data.

## Measuring performance

```{r profiling_packages_eval, echo = FALSE, ref.label = "profiling_packages_show", message = FALSE, results = "hide"}
```

Timings are measured using the `r BiocStyle::CRANpkg("microbenchmark")` package 
using `r times` repetitions (`times`). Memory allocations are reported using 
the `r BiocStyle::CRANpkg("profmem")` package.

```{r profiling_packages_show, echo = TRUE, message = FALSE}
library(microbenchmark)
library(profmem)
times <- 10
```
