# `colSums2()`

We compare the `DelayedArray::colSums()` method, which uses the 
block-processing algorithm, to the `DelayedMatrixStats::colSums2()` method, 
which uses tailored algorithms depending on the *seed* of the *DelayedMatrix* 
and to the `DelayedMatrixStats:::.colSums2()` method that acts directly on the 
seed of the *DelayedMatrix* object[^dotted_method]

[^dotted_method]: `DelayedMatrixStats:::.colSums2()` shouldn't be called by the user because it does not realise delayed operations. It is used here for demonstration purposes on a "*pristine*" DelayedMatrix to measure the additional overhead of S4 methods

## Pristine *DelayedMatrix*

```{r colSums2}
microbenchmark(DelayedArray::colSums(dense_matrix),
               DelayedMatrixStats::colSums2(dense_matrix),
               DelayedMatrixStats:::.colSums2(seed(dense_matrix)),
               times = times)
total(profmem(DelayedArray::colSums(dense_matrix)))
total(profmem(DelayedMatrixStats::colSums2(dense_matrix)))

microbenchmark(DelayedArray::colSums(sparse_matrix),
               DelayedMatrixStats::colSums2(sparse_matrix),
               DelayedMatrixStats:::.colSums2(seed(sparse_matrix)),
               Matrix::colSums(seed(sparse_matrix)), 
               times = times)
total(profmem(DelayedArray::colSums(sparse_matrix)))
total(profmem(DelayedMatrixStats::colSums2(sparse_matrix)))

microbenchmark(DelayedArray::colSums(rle_matrix),
               DelayedMatrixStats::colSums2(rle_matrix),
               DelayedMatrixStats:::.colSums2(seed(rle_matrix)), 
               times = times)

total(profmem(DelayedArray::colSums(rle_matrix)))
total(profmem(DelayedMatrixStats::colSums2(rle_matrix)))
```

## With row subsetting

```{r colSums2_with_rows}
i <- sample(nrow(dense_matrix), nrow(dense_matrix) / 10)
microbenchmark(DelayedArray::colSums(dense_matrix[i, ]),
               DelayedMatrixStats::colSums2(dense_matrix, rows = i),
               times = times)
total(profmem(DelayedArray::colSums(dense_matrix[i, ])))
total(profmem(DelayedMatrixStats::colSums2(dense_matrix, rows = i)))

microbenchmark(DelayedArray::colSums(sparse_matrix[i, ]),
               DelayedMatrixStats::colSums2(sparse_matrix, rows = i),
               times = times)
total(profmem(DelayedArray::colSums(sparse_matrix[i, ])))
total(profmem(DelayedMatrixStats::colSums2(sparse_matrix, rows = i)))

i <- sample(nrow(rle_matrix), nrow(rle_matrix) / 10)
microbenchmark(DelayedArray::colSums(rle_matrix[i, ]),
               DelayedMatrixStats::colSums2(rle_matrix, rows = i),
               times = times)
total(profmem(DelayedArray::colSums(rle_matrix[i, ])))
total(profmem(DelayedMatrixStats::colSums2(rle_matrix, rows = i)))
```

## With column subsetting

```{r colSums2_with_cols}
j <- sample(ncol(dense_matrix), ncol(dense_matrix) / 10)
microbenchmark(DelayedArray::colSums(dense_matrix[, j]),
               DelayedMatrixStats::colSums2(dense_matrix, cols = j),
               times = times)
total(profmem(DelayedArray::colSums(dense_matrix[, j])))
total(profmem(DelayedMatrixStats::colSums2(dense_matrix, cols = j)))

microbenchmark(DelayedArray::colSums(sparse_matrix[, j]),
               DelayedMatrixStats::colSums2(sparse_matrix, cols = j),
               times = times)
total(profmem(DelayedArray::colSums(sparse_matrix[, j])))
total(profmem(DelayedMatrixStats::colSums2(sparse_matrix, cols = j)))

j <- sample(ncol(rle_matrix), ncol(rle_matrix) / 2)
microbenchmark(DelayedArray::colSums(rle_matrix[, j]),
               DelayedMatrixStats::colSums2(rle_matrix, cols = j),
               times = times)
total(profmem(DelayedArray::colSums(rle_matrix[, j])))
total(profmem(DelayedMatrixStats::colSums2(rle_matrix, cols = j)))
```

## With row and column subsetting

```{r colSums2_with_rows_and_cols}
i <- sample(nrow(dense_matrix), nrow(dense_matrix) / 10)
j <- sample(ncol(dense_matrix), ncol(dense_matrix) / 10)
microbenchmark(DelayedArray::colSums(dense_matrix[i, j]),
               DelayedMatrixStats::colSums2(dense_matrix, rows = i, cols = j),
               times = times)
microbenchmark(DelayedArray::colSums(sparse_matrix[i, j]),
               DelayedMatrixStats::colSums2(sparse_matrix, rows = i, cols = j),
               times = times)
i <- sample(nrow(rle_matrix), nrow(rle_matrix) / 10)
j <- sample(ncol(rle_matrix), ncol(rle_matrix) / 2)
microbenchmark(DelayedArray::colSums(rle_matrix[i, j]),
               DelayedMatrixStats::colSums2(rle_matrix, rows = i, cols = j),
               times = times)
```

## With delayed ops

```{r colSums2_with_delayed_ops}
microbenchmark(DelayedArray::colSums(dense_matrix ^ 2),
               DelayedMatrixStats::colSums2(dense_matrix ^ 2),
               times = times)
microbenchmark(DelayedArray::colSums(sparse_matrix ^ 2),
               DelayedMatrixStats::colSums2(sparse_matrix ^ 2),
               times = times)
microbenchmark(DelayedArray::colSums(rle_matrix ^ 2),
               DelayedMatrixStats::colSums2(rle_matrix ^ 2),
               times = times)
```

## Summary
